---
title: "p8105_hw2_zc2691"
author: "Zhaohua Chunyu"
output: github_document
date: "2022-10-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(tidyverse)
library(readxl)
library(dplyr)
```

```{r}
subway_df = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                     col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>% 
  select(line:route11, entry:vending, ada, entrance_type) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
subway_df
```

```{r}
subway_df %>% 
  select(station_name, line) %>% 
  distinct
```

```{r}
subway_df %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

```{r}
subway_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "routenum",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select (line, station_name) %>% 
  distinct
```

```{r}
subway_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "routenum",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select (line, station_name) %>% 
  distinct
```

## Probelm 2
```{r}
mr_trash_df =  read_excel("data/Trash Wheel Collection Data.xlsx", sheet = 1) %>% 
  janitor::clean_names() %>% 
  select(dumpster:homes_powered) %>% 
  drop_na() %>% 
  mutate(year = as.character(year)) %>% 
  mutate(weight_tons = round(weight_tons, digit = 0)) %>% 
  mutate(weight_tons = as.integer(weight_tons)) %>% 
  mutate(id = "Mr.Trash Wheel") %>% 
  select(id, everything())
mr_trash_df

prof_trash_df =  read_excel("data/Trash Wheel Collection Data.xlsx", sheet = 2) %>% 
  janitor::clean_names() %>% 
  select(dumpster:homes_powered) %>% 
  drop_na() %>% 
  mutate(year = as.character(year)) %>% 
  mutate(weight_tons = round(weight_tons, digit = 0)) %>% 
  mutate(weight_tons = as.integer(weight_tons)) %>% 
  mutate(id = "Professor Trash Wheel") %>% 
  select(id, everything())
prof_trash_df
```

```{r}
trash_df = bind_rows(mr_trash_df, prof_trash_df) 
```

```{r}
summary(mr_trash_df) ## size of df: 486* 15
summary(prof_trash_df) ## size of df: 82 * 14
summary(trash_df) ## sizeof df: 568 * 15
mr_trash_df[1,] ##Examples of an observation of the dataframe
prof_trash_df[1,] ##Examples of an observation of the dataframe
```

```{r}
prof_trash_df %>% 
  select(weight_tons) %>% 
  pull(weight_tons) %>% 
  sum() ## Calculate the total weight of trash collected by Professor Trash Wheel

mr_trash_df %>% 
  select(year, sports_balls) %>% 
  filter(year == "2020") %>%
  pull(sports_balls) %>% 
  sum() ## Calculate the total number of sports balls collected by Mr. Trash Wheel in 2020
```

There are 486 observations in *mr_trash_df* dataset and 82 observations in *prof_trash_df* dataset Each table contains variables of dumpter number (`dumpster`), date of collection (`date`), amount of total litter(`weight_tons` & `volume_cubic_yards`) and different litter types (such as `plastic bottles`, `cigarette butts`, `sport balls`, and etc.) The *trash_df* combine the dumpsters data of *mr_trash_df* and *prof_trash_df* by `bind_rows`. The total weight of trash collected by Professor Trash Wheel was 166 tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 was 856. 

## Problem 3
```{r}
pols_mon_raw = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(prez_gop = replace(prez_gop, prez_gop == 1, "gop")) %>%
  mutate(prez_gop = replace(prez_gop, prez_gop == 0, "dem")) %>%
  mutate(prez_dem = replace(prez_dem, prez_dem == 1, "dem")) %>%
  mutate(prez_dem = replace(prez_gop, prez_gop == 0, "gop")) %>%
  pivot_longer(
    c(prez_dem, prez_gop),
    names_to = "party",
    values_to = "president") %>% 
  select(year,month, president, gov_gop:rep_gop, gov_dem:rep_dem)
  
rows = nrow(pols_mon_raw)
odd_rows = seq_len(rows) %% 2
pols_mon = pols_mon_raw[odd_rows == 1, ]
pols_mon

  
snp = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>% 
  separate(date, c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month]) %>% 
  select(year,month, close) 
snp
 
unemp = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemp_perc"
  ) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(month = replace(month, month == "jan", "January")) %>% 
  mutate(month = replace(month, month == "feb", "February")) %>% 
  mutate(month = replace(month, month == "mar", "March")) %>% 
  mutate(month = replace(month, month == "apr", "April")) %>%
  mutate(month = replace(month, month == "may", "May")) %>% 
  mutate(month = replace(month, month == "jun", "June")) %>% 
  mutate(month = replace(month, month == "jul", "July")) %>% 
  mutate(month= replace(month, month == "aug", "August")) %>% 
  mutate(month = replace(month, month == "sep", "September")) %>% 
  mutate(month = replace(month, month == "oct", "October")) %>% 
  mutate(month = replace(month, month == "nov", "November")) %>% 
  mutate(month = replace(month, month == "dec", "December")) 
unemp
```

```{r}
pols_mon_snp = left_join(pols_mon, snp, by = c("year", "month"))
pols_mon_snp

pols_mon_snp_unemp = left_join(pols_mon_snp, unemp, by = c("year", "month"))
pols_mon_snp_unemp
```

```{r}
summary(pols_mon) ## size of df: 822* 9
summary(snp) ## size of df: 787 * 3
summary(unemp) ## size of df: 816 * 3
summary(pols_mon_snp_unemp) ## size of df: 822* 11
```

*pols_mon* dataset contains 822 observations of 9 variables related to the number of national politicians who are democratic or republican at any given time: `year` & `month`(year and month of the count), `president`(the indicator of whether the president was democratic or republican), `gov_gop`(the number of republican governors on the associated date), `sen_gop`(the number of republican senators on the associated date), `rep_gop`(the number of republican representatives on the associated date), etc.

*snp* contains 787 observations of 3 variables related to Standard & Poor’s stock market index (S&P):`year` & `month`, `close`(the closing values of the S&P stock index on the associated date). The max closing value is 2107.39; the min is 17.05; and the mean is 474.89.

*unemp* contains 816 observations of 3 variables:`year` & `month`, `unemp_perc`(percentage of unemployment in each month of the associated year). The max unemployment percentage is 10.80%; the min is 2.50%; and the mean is 5.83. 

The resulting dataset *pols_mon_snp_unemp* contains 816 observations of 11 variables obtained from the former three tables combined together. The range of year is from 1969 to 2015. The key variables include `year` & `month`, `president`, `gov_gop`, `sen_gop`, `rep_gop`, `close`, and `unemp_perc`. The mean close value is 241.1 and the mean unemployment percentage is 5.83. 



