---
title: "p8105_hw2_zc2691"
output: html_document
date: "2022-10-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(tidyverse)
library(readxl)
library(dplyr)
```

```{r}
subway_df = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                     col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>% 
  select(line:route11, entry:vending, ada, entrance_type) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

```{r}
subway_df %>% 
  select(station_name, line) %>% 
  distinct
```

```{r}
subway_df %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

```{r}
subway_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "routenum",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select (line, station_name) %>% 
  distinct
```

```{r}
subway_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "routenum",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select (line, station_name) %>% 
  distinct
```

## Probelm 2
```{r}
mr_trash_df =  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 1) %>% 
  janitor::clean_names() %>% 
  select(dumpster:homes_powered) %>% 
  drop_na() %>% 
  mutate(dumpster = as.character(dumpster)) %>% 
  mutate(weight_tons = round(weight_tons, digit = 0)) %>% 
  mutate(weight_tons = as.integer(weight_tons)) %>% 
  mutate(id = "Mr.Trash Wheel")  

prof_trash_df =  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 2) %>% 
  janitor::clean_names() %>% 
  select(dumpster:homes_powered) %>% 
  drop_na() %>% 
  mutate(dumpster = as.character(dumpster)) %>% 
  mutate(weight_tons = round(weight_tons, digit = 0)) %>% 
  mutate(weight_tons = as.integer(weight_tons)) %>% 
  mutate(id = "Professor Trash Wheel") 
```

```{r}
trash_df = full_join(mr_trash_df, prof_trash_df, by = "id")
```

```{r}
summary(mr_trash_df)
summary(prof_trash_df)
mr_trash_df[1,]
prof_trash_df[1,]
```

```{r}
prof_trash_df %>% 
  select(weight_tons) %>% 
  mutate(weight_tons_total = sum(weight_tons))
  sum(prof_trash_df$weight_tons)

mr_trash_df %>% 
  select(year, sports_balls) %>% 
  filter(year == "2020") %>% 
  mutate(sports_balls_2020_total = sum(sports_balls)) 
```

There are 453 observations in "Mr Trash Wheel" table and 71 observations in "Professor Wheel" table. Each table contains variables of dumpter number (`dumpster`), date of collection (`date`), amount of total litter(`weight_tons` & `volume_cubic_yards`) and different litter types (such as `plastic bottles`, `cigarette butts`, `sport balls`, and etc.) The total weight of trash collected by Professor Trash Wheel was 140 tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 was 856. 








